---
title: "Reproducible Research P1"
author: "Mauricio Rueda"
date: "19 de abril de 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Settings and libraries 
```{r setting up}
Sys.setlocale("LC_ALL","English")
setwd("~/GitHub/RepData_PeerAssessment1")
library(lattice)
library(ggplot2)
```

### tasks list

1. Load the data (i.e. read.csv()\color{red}{\verb|read.csv()|}read.csv())
```{r load data}
datos <- read.csv("activity.csv",stringsAsFactors = FALSE)
```

1.1. Process/transform the data (if necessary) into a format suitable for your analysis
```{r process data}
datos$date <- as.Date(datos$date)

datos_hist <- datos[!is.na(datos$steps),]
```

2.Histogram of the total number of steps taken each day

```{r task2}
tabla.sum <- aggregate(datos_hist[,1], list(datos_hist$date), sum)
colnames(tabla.sum) <- c("date","steps")
ggplot(tabla.sum,aes(x=date,y=steps))+geom_bar(stat='identity')
```

3.Mean and median number of steps taken each day
```{r task3}
tabla.mean <- aggregate(datos_hist[,1], list(datos_hist$date), mean)

colnames(tabla.mean) <- c("date","mean")

print(tabla.mean,type="html")

tabla.median <- aggregate(datos_hist[,1], list(datos_hist$date), median)

colnames(tabla.median) <- c("date","median")

print(tabla.median,type="html")

```

4.Time series plot of the average number of steps taken
```{r task4}
tabla.five <- aggregate(datos_hist[,1], list(datos_hist$interval), mean)

colnames(tabla.five) <- c("interval","mean")
```

```{r figura1,fig.showtext="fig1"}
plot(tabla.five$interval,tabla.five$mean, type="l", col="green", 
     xlab="5-minute interval", ylab="AVG Steps", main="average daily activity pattern")
```



5.The 5-minute interval that, on average, contains the maximum number of steps
```{r task5}
#maximum number of steps
tabla.five[which.max(tabla.five$mean),]
```

6.Code to describe and show a strategy for imputing missing data

6.1 Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NA\color{red}{\verb|NA|}NAs)

```{r task61}
print(colSums(is.na(datos)),type="html")
```

6.2 replace de mising steps data with the mean of steps for its own interval

```{r task62}
for(i in 1:nrow(datos)){
  if(is.na(datos[i,1])){
    intervalo <- datos[i,3]
    mean_steps <- round(mean(datos[datos$interval == intervalo,1],na.rm = TRUE))
    datos[i,1] <- mean_steps
  }
}
```

7. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.
```{r task7}
ggplot(datos,aes(x=date,y=steps))+geom_bar(stat='identity')

tabla.mean_new <- aggregate(datos[,1], list(datos$date), mean)

colnames(tabla.mean_new) <- c("date","mean")

tabla.median_new <- aggregate(datos[,1], list(datos$date), median)

colnames(tabla.median_new) <- c("date","median")
```


As you can see the missing data was related with missing days. So now the analisys looks complete.

8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

8.1 Create a new factor variable in the dataset with two levels (weekday and weekend) indicating whether a given date is a weekday or weekend day.

```{r task8}
datos_last <- datos

weekends <- c("Saturday","Sunday")

datos_last$day_type <- apply(datos_last, 1, FUN = function(x){
  if(weekdays(as.Date(x[2])) %in% weekends){
    "weekend"   
  }
  else
    "weekday"
})

datos_last$day_type <- as.factor(datos_last$day_type)

tabla.by_type <- aggregate(datos_last[,1], list(datos_last$interval,datos_last$day_type), sum)

colnames(tabla.by_type) <- c("interval","day_type","steps")
```

8.2 Make a panel plot containing a time series plot (i.e. type = "l"\color{red}{\verb|type = "l"|}type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r task9}
#maximum number of steps
xyplot(steps~interval|day_type,data=tabla.by_type,type="l",scales=list(y=list(relation="free")),layout=c(1,2))
```

